---
title: "Lab 5: Factors in Visualizations"
author: "Elizabeth Scholtes"
format:
  html:
    self-contained: true
editor: visual
---

# Question 1

```{r}
#| message: false
library(tidyverse)
```

```{r}
#| message: false
surveys <- read_csv(here::here("supporting_artifacts",
                               "learning_targets",
                               "Lab5",
                               "surveys.csv")) 
```

```{r}
ggplot(data = surveys, 
       mapping = aes(x = weight, 
                     y = species)
       ) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(size = .25, 
              color = "Purple", 
              alpha = .20) +
  xlab("Weight of Species (in grams)") +
  ylab("Type of Species")
```

# Question 2

Reorder the box plots so the weights go in descending order.

Moving the Y-axis to the top of the plot.

```{r}
surveys <- surveys |>
  mutate(
    species = fct_reorder(species,
                          weight,
                          .desc = FALSE)
    )
ggplot(data = surveys, 
       mapping = aes(x = weight,
                     y = species)
       ) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(size = .25, 
              color = "Purple", 
              alpha = .20) +
  labs(x = "Weight of Species (in grams)",
       y = "") +
  ggtitle("Type of Species")
```

# Part 2: Time Series Plot

# Question 3

Weights over time from original surveys data

```{r}
ggplot(data = surveys,
       mapping = aes(x = year,
                     y = weight,
                     color = genus)) +
  geom_point()
```

# Question 4

Mean weights by year and genus over time

```{r}
weightavg <- surveys |>
  group_by(genus, year) |>
  summarize(across(.cols = weight,
                   mean,
                   na.rm = TRUE),
            .groups = "keep"
            )
ggplot(data = weightavg,
       mapping = aes(x = year,
                     y = weight,
                     color = genus)) +
  geom_point() +
  geom_line()
```

# Question 5

Reorder by weight and rename legend title and move legend info to plot title

```{r}
weightavg <- surveys |>
  group_by(genus, year) |>
  summarize(across(.cols = weight,
                   mean,
                   na.rm = TRUE),
            .groups = "keep"
            )

ggplot(data = weightavg,
       mapping = aes(x = year,
                     y = weight,
                     color = fct_reorder2(genus,
                         year,
                         weight,
                         .desc = TRUE))) +
  geom_point() +
  geom_line() +
  labs(x = "Year",
       y = "",
       color = "Species") +
  ggtitle("Weight")
```

# Part 3: Captures Over the Week

# Question 6

Rodents captured each day of the week

```{r}
surveys <- surveys |>
  filter(
    day_of_week != is.na(day_of_week)
  )
ggplot(data = surveys,
       mapping = aes(x = day_of_week)) +
  geom_bar() +
  labs(y = "") +
  ggtitle("Number of Rodents Captured")
```

# Question 7

change order of day of week to Mon-Sun

```{r}
surveys <- surveys|>
  mutate(
    day_of_week = fct_relevel(day_of_week,
                              levels = c("Mon",
                                         "Tue",
                                         "Wed",
                                         "Thu",
                                         "Fri",
                                         "Sat",
                                         "Sun"))
  )

ggplot(data = surveys,
       mapping = aes(x = day_of_week)) +
  geom_bar() +
  labs(y = "") +
  ggtitle("Number of Rodents Captured")
```

# Question 8

Create weekday and weekend levels

```{r}
week <- surveys |>
  mutate(
    weekday = fct_collapse(day_of_week,
                           weekend = c("Sat", "Sun"),
                           weekday = c("Mon",
                                       "Tue",
                                       "Wed",
                                       "Thu",
                                       "Fri"))
    )

ggplot(data = week,
       mapping = aes(x = weekday)) +
  geom_bar() +
  labs(x = "Day of the Week",
       y = "") +
  ggtitle("Number of Captures")
```
