---
title: "Challenge 9: Baby Names"
author: "Elizabeth Scholtes"
format:
  html:
    self-contained: true
editor: visual
---

```{r}
#| message: false
library(tidyverse)
library(scales)
names <- read_csv(here::here("supporting_artifacts", "learning_targets", "Challenge 9", "StateNames_A.csv"))
```

```{r}
# Rename column as Sex
names <- rename(names, "Sex" = "Gender")
```

# Challenge: Question 1

```{r}
#| message: false
library(DT)
```

```{r}
#| warning: false
datatable(names)
```

```{r}
#| message: false
allison_names <- names |>
  group_by(State, Sex) |>
  filter(Name == "Allison") |>
  summarize(Count = sum(Count)) |>
  pivot_wider(names_from = Sex,
              values_from = Count) |>
  mutate(
    M = coalesce(M, 0)
  )
allison_names
```

```{r}
allison_F_names <- names |>
  filter(Name == "Allison",
         Sex == "F")
```

```{r}
# Create data frame that can be used in future questions
names_by_year <- allison_F_names |>
  group_by(Year) |>
  summarize(Count = sum(Count))

names_by_year |>
  ggplot(mapping = aes(x = Year,
                       y = Count)) +
  geom_point() +
  geom_line() +
  labs(title = "Number of people assigned female at birth with the name Allison")
```

```{r}
allison_lm <- names_by_year |>
  lm(Count ~ Year, data = _)
allison_lm
```

```{r}
names_by_year |>
  ggplot(aes(x = Year,
             y = Count)) +
  geom_point() +
  stat_smooth(method = "lm")
```

Estimated Regression Equation:

**Count(hat) = 209689.8 - 101.5\*Year**

```{r}
allison_lm |>
  broom::augment() |>
  ggplot(mapping = aes(x = .resid,
                       y = Count)) +
  geom_point() +
  labs(x = "Residuals")
```

It looks like there is a pattern with the residuals in that the residual value gets higher as the number of "Allisons" per year gets higher. This is not ideal because that means that the residuals are not random and do not necessarily satisfy linear model conditions

From this model, we can conclude that the name Allison is becoming less and less popular. Maybe in terms of how often the name is used it's less "cool", but that's not necessarily true!

```{r}
#| message: false
al_names <- data.frame(Name = c("Allan", "Alan", "Allen"))

allan_names <- names |>
  semi_join(al_names,
            names,
            by = "Name") |>
  filter(Sex == "M")

allan_names |>
  group_by(Year, Name) |>
  summarize(Count = sum(Count))|>
  ggplot(mapping = aes(x = Year,
                       y = Count,
                       color = Name)) +
  geom_point() +
  geom_line() +
  labs(title = "Popularity of names 'Alan', 'Allan', and 'Allen'")
```

```{r}
statenames <- allan_names |>
  pivot_wider(names_from = Name, values_from = Count) |>
  filter(Year == 2000,
         State == "CA" | State == "PA") |>
  select(State, Allen, Alan, Allan)
statenames
```

```{r}
percents <- statenames |>
  rowwise()|>
  mutate(
    total_sum = sum(c_across(Allen : Allan)),
    Allen = percent( (Allen / total_sum), accuracy = .01),
    Alan = percent( (Alan / total_sum), accuracy = .01),
    Allan = percent( (Allan / total_sum), accuracy = .01) 
    ) |>
  select(State:Allan)
percents
```

# Challenge: Question 2

```{r}
#| message: false
library(kableExtra)
```

```{r}
# Table of Allisons by State
knitr::kable(allison_names,
             format = "html",
             col.names = c("State",
                           "Sex assigned at Birth",
                           "Number of People with Name Allison"),
             booktabs = TRUE) |>
  kable_styling(latex_options = "scale_down")
```

```{r}
# Allan names raw counts
knitr::kable(statenames,
              format = "html",
             caption = "Number of 'Alans' by State",
             booktabs = TRUE) |>
  add_header_above(c("Location" = 1, "Name" = 3))
```

```{r}
# Allan names by percent
knitr::kable(percents,
             format = "html",
             caption = "Percentage of 'Alans' by State",
             booktabs = TRUE) |>
  kable_styling(font_size = 18)
```
